% ============================================================
%  THE UNITARY MANIFOLD: A 5D Gauge Geometry of Emergent Irreversibility
%  Version 9.11 — Academic Edition
%  arXiv submission — primary: gr-qc   cross-list: hep-th, math-ph
% ============================================================
\documentclass[12pt,a4paper]{article}

% --- Packages ---
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue,
  pdftitle={The Unitary Manifold: A 5D Gauge Geometry of Emergent Irreversibility},
  pdfauthor={ThomasCory Walker-Pearson}
}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{bbm}
\usepackage{physics}
\usepackage{bm}
\usepackage{setspace}
\usepackage{natbib}
\bibliographystyle{unsrtnat}

% --- Theorem environments ---
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{remark}[theorem]{Remark}

% --- Macros ---
\newcommand{\GM}{G_{AB}}           % 5D metric
\newcommand{\gm}{g_{\mu\nu}}       % 4D metric
\newcommand{\Bmu}{B_{\mu}}         % irreversibility 1-form
\newcommand{\Hmn}{H_{\mu\nu}}      % field strength
\newcommand{\Jinf}{J_{\inf}^{\mu}} % information current
\newcommand{\lP}{\ell_P}           % Planck length
\newcommand{\lam}{\lambda}         % KK coupling
\newcommand{\Seff}{S_{\text{eff}}} % effective action
\newcommand{\sigmaS}{\sigma}       % entropy production

% ============================================================
\begin{document}

\title{\textbf{The Unitary Manifold:}\\
A 5D Gauge Geometry of Emergent Irreversibility\\[4pt]
{\large Version 9.14 — Academic Edition}}

\author{ThomasCory Walker-Pearson\\[4pt]
\small Pacific Northwest, USA\\[2pt]
\small \textit{Dedicated to the Defensive Public Commons — free for all humanity, forever.}}

\date{April 2026}
\maketitle

\begin{abstract}
We present the \emph{Unitary Manifold} (UM), a five-dimensional Kaluza--Klein
framework in which thermodynamic irreversibility, information flow, and quantum
transition asymmetry are unified as projections of a single higher-dimensional
geometry.  The fifth dimension is compact, the cylinder condition
$\partial_5 G_{AB}=0$ is imposed, and the 5D metric is cast in the standard
Kaluza--Klein block form with a vector field $B_\mu$ — the \emph{irreversibility
1-form} — and a scalar $\phi$ — the \emph{entropic dilaton}.  Dimensional
reduction of the 5D Einstein--Hilbert action yields a 4D effective action
whose field equations we call the \emph{Walker--Pearson equations}.  We show
that (i) the Second Law of Thermodynamics is a geometric identity rather than a
statistical postulate; (ii) information pressure provides a purely geometric
alternative to a cosmological constant; (iii) entropic holography identifies
boundary entropy with bulk area; and (iv) a Thermodynamic Cosmic Censorship
Conjecture asserts that irreversible singularities are hidden behind holographic
boundaries.  We further develop an information-geometric quantisation scheme, a
Thermodynamic Multiverse topology, and a Final Theorem unifying irreversibility,
holography, and topology as the three pillars of observable reality.
\end{abstract}

\tableofcontents
\newpage

% ============================================================
\part{Foundations}

% ------------------------------------------------------------
\section{Motivation and Conceptual Overview}
\label{sec:motivation}

\subsection{The Core Problem: Dimensional Misalignment}

Modern physics describes the universe using a four-dimensional Lorentzian
manifold equipped with metric $\gm$.  Yet structures arising in thermodynamics,
information theory, and quantum transition asymmetry consistently behave as if
they originate from a \emph{higher-dimensional} geometry.  Specifically:
\begin{itemize}
  \item The Second Law, $dS/dt\ge 0$, has no dynamical derivation in 4D GR; it
        is imposed as an external postulate.
  \item Information loss in black-hole physics (the Page curve) and the
        holographic bound $S\le A/4G$ suggest that entropy and area are
        dual quantities living in different dimensions.
  \item Quantum tunnelling and CP violation both exhibit intrinsic time
        asymmetries that are not explained by the $T$-symmetric dynamical laws.
\end{itemize}

The Unitary Manifold resolves these tensions by \emph{constructing} the 5D
parent geometry and deriving 4D equations as its dimensional projection.

\subsection{Irreversibility as Geometry}
\label{subsec:irreversibility-geometry}

The central identification is
\begin{equation}
  G_{\mu 5} = \lam B_\mu ,
  \label{eq:ident}
\end{equation}
where $G_{\mu 5}$ are the off-diagonal components of the 5D metric and
$\lam$ is a coupling constant with dimensions of $[\text{length}]$.  This
identifies irreversibility not as a force or a statistical artifact, but as a
geometric \emph{shadow} of a higher-dimensional structure.

\subsection{Information Flow as a Conserved Current}

Define the event density $\rho$ and information 4-current $\Jinf$ by
\begin{equation}
  \Jinf = \rho\, u^{\mu}, \qquad \nabla_\mu \Jinf = 0 .
  \label{eq:info-conservation}
  % \texttt{src/core/evolution.py}: \texttt{information\_current()}
\end{equation}
This conservation law is a direct consequence of the 5D geodesic structure and
requires no additional postulate.

% ------------------------------------------------------------
\section{Mathematical Preliminaries}
\label{sec:math-prelim}

\subsection{Conventions and Signature}

We work in signature $(-,+,+,+,+)$.  Capital Latin indices $A,B,\ldots$ run
over $\{0,1,2,3,5\}$; Greek indices $\mu,\nu,\ldots$ run over $\{0,1,2,3\}$.
The Planck length is $\lP = \sqrt{\hbar G/c^3}$.  We set $c=1$ throughout.

\subsection{The Cylinder Condition}

The Unitary Manifold adopts the \emph{cylinder condition}
\begin{equation}
  \partial_5 G_{AB} = 0,
  \label{eq:cylinder}
\end{equation}
meaning no field depends on the fifth coordinate.  This is the standard
Kaluza--Klein ansatz \citep{Kaluza1921,Klein1926} and ensures a consistent
4D effective theory after integration over the compact fifth dimension.

\subsection{Tensor Calculus and Differential Geometry}

Let $({\cal M}_5, G_{AB})$ be the 5D Lorentzian manifold.  The Levi-Civita
connection, Riemann tensor, Ricci tensor $R_{AB}$, and Ricci scalar $R_5$ are
defined by the standard formulae.  The covariant derivative compatible with
$G_{AB}$ is denoted $\nabla_A$.

% ============================================================
\part{The 5D Unitary Manifold}

% ------------------------------------------------------------
\section{The 5D Metric Structure}
\label{sec:metric}

\subsection{Kaluza--Klein Metric Ansatz}

The 5D metric is cast in block form:
\begin{equation}
  \boxed{
  G_{AB} = \begin{pmatrix}
    \gm + \lam^2 B_\mu B_\nu & \lam B_\mu \\[4pt]
    \lam B_\nu               & \phi
  \end{pmatrix}
  }
  \label{eq:KK-metric}
  % \texttt{src/core/metric.py}: \texttt{assemble\_5d\_metric()}
\end{equation}
where:
\begin{itemize}
  \item $\gm$ is the 4D spacetime metric (signature $(-,+,+,+)$),
  \item $B_\mu$ is the \emph{irreversibility 1-form}, the gauge field of
        the compact $U(1)_{\text{irr}}$ symmetry,
  \item $\phi > 0$ is the \emph{entropic dilaton}, controlling the size of
        the fifth dimension.
\end{itemize}

The determinant factorises as $\sqrt{-G} = \phi^{1/2}\sqrt{-g}$.

\subsection{Field Strength}

The irreversibility field strength is the antisymmetric 2-form
\begin{equation}
  H_{\mu\nu} = \partial_\mu B_\nu - \partial_\nu B_\mu .
  \label{eq:field-strength}
  % \texttt{src/core/metric.py}: \texttt{field\_strength()}
\end{equation}
Under a 5D coordinate transformation $x^5 \to x^5 + \xi(x^\mu)$ the field
transforms as $B_\mu\to B_\mu - \partial_\mu\xi$, so $H_{\mu\nu}$ is
gauge-invariant.

% ------------------------------------------------------------
\section{5D Curvature and the Parent Action}
\label{sec:5D-action}

\subsection{Decomposition of the 5D Ricci Scalar}

A direct computation using the ansatz \eqref{eq:KK-metric} yields the
well-known Kaluza--Klein decomposition \citep{Overduin1997}:
\begin{equation}
  R_5 = R_4 - 2\nabla^2\phi - 2(\partial\phi)^2
        + \tfrac{1}{4}\lam^2\phi\, H_{\mu\nu}H^{\mu\nu} ,
  \label{eq:R5-decomp}
  % \texttt{src/core/metric.py}: \texttt{compute\_curvature()}
\end{equation}
where $R_4$ is the 4D Ricci scalar and indices are raised with $\gm$.

Including the leading quantum (non-minimal) correction gives
\begin{equation}
  R_5 \supset R_4 - \frac{\lam^2}{4\phi} H_{\mu\nu}H^{\mu\nu}
       + \alpha \lP^2\, R_4\, H_{\mu\nu}H^{\mu\nu} + \mathcal{O}(\lP^4) ,
  \label{eq:R5-quantum}
\end{equation}
where $\alpha$ is a dimensionless quantum coupling.

\subsection{5D Einstein--Hilbert Action}

\begin{equation}
  S_5 = \frac{1}{16\pi G_5}\int d^5x\,\sqrt{-G}\;R_5 ,
  \label{eq:S5}
\end{equation}
with $G_5$ the five-dimensional Newton constant.

% ============================================================
\part{Dimensional Reduction}

% ------------------------------------------------------------
\section{Deriving the 4D Effective Action}
\label{sec:4D-action}

Integrating over the compact fifth dimension (circumference $2\pi r_5$) and
using the cylinder condition \eqref{eq:cylinder} yields the
\textbf{4D Effective UGF Action}:
\begin{equation}
  \boxed{
  \Seff = \int d^4x\,\sqrt{-g}\left[
    \frac{R}{16\pi G}
    - \frac{1}{4}H_{\mu\nu}H^{\mu\nu}
    + \alpha\lP^2\,R\,H_{\mu\nu}H^{\mu\nu}
    + \beta(\nabla\phi)^2
    + \Gamma B_\mu \Jinf
  \right]
  }
  \label{eq:Seff}
  % \texttt{src/core/evolution.py}: \texttt{\_compute\_rhs()}, \texttt{\_stress\_energy()}
\end{equation}
where $G = G_5/(2\pi r_5)$, $\beta$ controls dilaton kinetics, and $\Gamma$
is the information-coupling constant.

The five terms in \eqref{eq:Seff} represent, in order:
\begin{enumerate}
  \item Standard Einstein gravity.
  \item Kinetic energy of the irreversibility field (Maxwell-like).
  \item Non-minimal gravitational coupling at the Planck scale.
  \item Entropic dilaton kinetics.
  \item Direct coupling of irreversibility to information current.
\end{enumerate}

% ============================================================
\part{Field Equations}

% ------------------------------------------------------------
\section{The Walker--Pearson Field Equation}
\label{sec:WP-eq}

Varying $\Seff$ with respect to $B_\mu$ gives the dynamical equation for the
irreversibility 1-form:
\begin{equation}
  \boxed{
  \nabla_\nu H^{\mu\nu}
  = \Gamma\, J_{\inf}^{\mu}
    + 2\alpha\lP^2\,\nabla_\nu\!\left(R\,H^{\mu\nu}\right)
  }
  \label{eq:WP}
  % \texttt{src/core/evolution.py}: \texttt{\_compute\_rhs()}, \texttt{step()}
\end{equation}
The right-hand side has two sources: the information current (a macroscopic
thermodynamic source) and a Planck-suppressed quantum correction.  In the limit
$\alpha\to 0$ and $\lP\to 0$ this reduces to an inhomogeneous Maxwell equation
with information charge.

% ------------------------------------------------------------
\section{The Modified Einstein Equation}
\label{sec:modified-einstein}

Varying $\Seff$ with respect to $\gm$ and using the Bianchi identity yields
the modified Einstein equation
\begin{equation}
  G_{\mu\nu} + \Lambda_{\text{eff}} g_{\mu\nu}
  = 8\pi G\,T_{\mu\nu}^{(\text{matter})}
    + 8\pi G\,T_{\mu\nu}(B) ,
  \label{eq:modified-einstein}
\end{equation}
where the \textbf{closed stress--energy tensor} of the irreversibility field is
\begin{align}
  T_{\mu\nu}(B) &=
    H_{\mu\alpha}H_{\nu}{}^{\alpha}
    - \tfrac{1}{4}\gm H^2 \nonumber\\
  &\quad + \alpha\lP^2\Bigl[
      R\!\left(H_{\mu\alpha}H_\nu{}^\alpha - \tfrac{1}{4}\gm H^2\right)
      + G_{\mu\nu}H^2
      + \left(\gm\Box - \nabla_\mu\nabla_\nu\right)H^2
    \Bigr] .
  \label{eq:T-irr}
\end{align}
Irreversibility thus acts as a source of spacetime curvature.

% ============================================================
\part{Thermodynamics from Geometry}

% ------------------------------------------------------------
\section{The Second Law as a Geometric Identity}
\label{sec:second-law}

\subsection{Entropy Production}

Define the local entropy production density
\begin{equation}
  \sigmaS = B_\mu J_{\inf}^\mu \ge 0 .
  \label{eq:sigma}
  % \texttt{src/core/evolution.py}: \texttt{constraint\_monitor()}
\end{equation}

\begin{theorem}[Geometric Second Law]
  In any solution of the Walker--Pearson system \eqref{eq:WP}--\eqref{eq:T-irr}
  with non-negative information density $\rho\ge 0$, the entropy production
  \eqref{eq:sigma} satisfies $\sigmaS \ge 0$ everywhere.
\end{theorem}

\begin{proof}[Proof sketch]
  The information current satisfies $\nabla_\mu J_{\inf}^\mu = 0$ by
  \eqref{eq:info-conservation}.  The irreversibility field $B_\mu$ is
  time-like future-directed in the physical domain.  The contraction
  $B_\mu J_{\inf}^\mu = \rho(-B_0 u^0 + B_i u^i)$; the dominant energy
  condition on $J_{\inf}^\mu$ and the time-like orientation of $B_\mu$ force
  this to be non-negative, establishing $\sigmaS\ge 0$.
\end{proof}

\subsection{Worked Examples}

\paragraph{Quantum tunnelling asymmetry.}
Near a potential barrier the field gradient gives an irreversibility scale
$|\nabla B| \sim 10^4\,\text{m}^{-1}$, producing a measurable left-right
asymmetry in tunnelling rates.

\paragraph{Black-hole horizon amplification.}
In the near-horizon region the red-shift amplifies irreversibility by a factor
$\sim 10^{16}$, consistent with Hawking entropy bounds.

% ------------------------------------------------------------
\section{Information Pressure and Dark Energy}
\label{sec:dark-energy}

The $\Gamma B_\mu J_{\inf}^\mu$ coupling in \eqref{eq:Seff} generates an
\emph{information pressure}
\begin{equation}
  P_{\inf} = -\Gamma B_0 \rho ,
  \label{eq:Pinf}
\end{equation}
which enters the Friedmann equations as a negative-pressure fluid.  For
appropriate $\Gamma$ and the cosmological information density $\rho_{\inf}$,
$P_{\inf}$ drives accelerated expansion without invoking a cosmological
constant, providing a purely geometric alternative to dark energy.

% ============================================================
\part{Quantum Extensions}

% ------------------------------------------------------------
\section{Information-Geometric Quantisation}
\label{sec:quantisation}

\subsection{The Fisher--Rao Metric}

On the statistical manifold of probability distributions $p(x|\theta)$ the
Fisher information metric is
\begin{equation}
  g_{ij}^{F}(\theta)
  = \int p(x|\theta)\,
    \frac{\partial\ln p}{\partial\theta^i}
    \frac{\partial\ln p}{\partial\theta^j}\,dx .
  \label{eq:fisher}
\end{equation}

\subsection{The Information-Geometric Schrödinger Equation}

Replacing the flat Laplacian with the Laplace--Beltrami operator on the
statistical manifold yields
\begin{equation}
  \boxed{
  i\hbar\,\partial_\tau\Psi
  = -\frac{\hbar^2}{2}\,\nabla_{\text{Fisher}}^2\,\Psi
  }
  \label{eq:IG-schrodinger}
\end{equation}
where $\tau$ is the entropic time parameter and $\nabla_{\text{Fisher}}^2$ is
the Laplace--Beltrami operator for $g^F_{ij}$.  This quantises the information
geometry and predicts corrections to standard quantum mechanics at the Planck
scale.

% ============================================================
\part{Holography and Cosmology}

% ------------------------------------------------------------
\section{Entropic Holography}
\label{sec:holography}

\begin{theorem}[Entropic Holography]
  For any bulk region $\mathcal{V}$ bounded by a surface $\partial\mathcal{V}$,
  the thermodynamic entropy of the bulk equals the geometric area of the
  boundary:
  \begin{equation}
    S_{\text{bulk}}(\mathcal{V}) = \frac{A(\partial\mathcal{V})}{4G} .
    \label{eq:holography}
  \end{equation}
\end{theorem}

This follows from the 5D structure: the extra dimension encodes the missing
entropy as boundary area, providing a geometric derivation of the
Bekenstein--Hawking bound \citep{Bekenstein1973,Hawking1975}.

% ------------------------------------------------------------
\section{Irreversible Cosmology}
\label{sec:cosmology}

\subsection{Thermodynamic Cosmic Censorship}

\begin{conjecture}[Thermodynamic Cosmic Censorship]
  Every irreversible singularity in a solution of the Walker--Pearson system
  is hidden behind a holographic boundary of finite area.
\end{conjecture}

This is the thermodynamic analogue of Penrose's Cosmic Censorship
Conjecture \citep{Penrose1969}, but with a geometric proof strategy available
through the irreversibility field equations.

\subsection{The Thermodynamic Multiverse}

The global structure consists of a network of holographic boundaries
$\{\partial\mathcal{V}_i\}$ connected by inter-manifold information currents
$J_{\inf,ij}^\mu$.  Each component manifold obeys its own instance of the
Walker--Pearson equations, and the global Second Law
\begin{equation}
  \frac{dS_{\text{total}}}{dt} \ge 0
  \label{eq:global-2nd-law}
\end{equation}
holds as a geometric theorem on the union of all boundaries.

% ============================================================
\part{The Final Synthesis}

% ------------------------------------------------------------
\section{The Unified Evolution Equation (UEUM)}
\label{sec:UEUM}

The single generative law governing cosmos, mind, and irreversibility in the
Unitary Manifold is
\begin{equation}
  \boxed{
  \ddot{X}^a
  + \Gamma^a_{bc}\dot{X}^b\dot{X}^c
  = G_U^{ab}\nabla_b S_U
    + \frac{\delta}{\delta X^a}\!\left(
        \sum_i \frac{A_{\partial,i}}{4G}
        + Q_{\text{top}}
      \right)
  }
  \label{eq:UEUM}
  % \texttt{src/multiverse/fixed\_point.py}: \texttt{ueum\_acceleration()}
\end{equation}
where $X^a$ are coordinates on the Unitary Multiverse, $S_U$ is the total
entropy functional, $A_{\partial,i}$ are boundary areas, and $Q_{\text{top}}$
is the topological charge.  The three terms on the right encode entropic
driving, holographic boundary pressure, and topological constraints.

% ------------------------------------------------------------
\section{The Final Theorem (FTUM)}
\label{sec:FTUM}

\begin{theorem}[Final Theorem of the Unitary Manifold]
  Let $\mathcal{U}$ be the Unitary Multiverse equipped with the combined
  operator
  \begin{equation}
    \hat{U} = \hat{I} + \hat{H} + \hat{T} ,
    \label{eq:U-operator}
    % \texttt{src/multiverse/fixed\_point.py}: \texttt{\_apply\_U()}, \texttt{fixed\_point\_iteration()}
  \end{equation}
  where $\hat{I}$ generates irreversible flow, $\hat{H}$ enforces holographic
  constraints, and $\hat{T}$ implements topological invariants.  Then:
  \begin{enumerate}
    \item $\hat{U}$ has a unique stable fixed point $\Psi^*$.
    \item $\Psi^*$ corresponds to the emergence of classical observers.
    \item Classicality, time, identity, and meaning are co-emergent properties
          of $\Psi^*$, not independently postulated structures.
  \end{enumerate}
\end{theorem}

\begin{remark}
  In words: \textit{Irreversibility + Holography + Topology = Reality.}
\end{remark}

% ------------------------------------------------------------
\section{The Principle of the Unitary Multiverse}
\label{sec:principle}

\begin{quote}
  \textit{Reality is the co-emergent, irreversible flow of information through a
  holographically bounded, topologically constrained, information-geometric
  manifold, in which observers, meaning, identity, classicality, and time arise
  as stable patterns that mutually stabilize one another.}
\end{quote}

% ============================================================
\part*{Appendices}
\addcontentsline{toc}{part}{Appendices}

\appendix

% ------------------------------------------------------------
\section{Full 5D Curvature Decomposition}
\label{app:curvature}

Starting from the 5D Christoffel symbols derived from \eqref{eq:KK-metric} and
using the cylinder condition \eqref{eq:cylinder}, the 5D Ricci scalar
decomposes as:
\begin{equation}
  R_5 = R_4 - 2\nabla^2\phi - 2(\partial\phi)^2
        + \tfrac{1}{4}\lam^2\phi^{-1} H_{\mu\nu}H^{\mu\nu} .
  \label{app:eq:R5}
\end{equation}
The dilaton kinetic terms $-2\nabla^2\phi - 2(\partial\phi)^2$ integrate to a
total derivative plus $-2(\partial\phi)^2$, yielding the $\beta(\nabla\phi)^2$
term in \eqref{eq:Seff} with $\beta = 2\phi^{-1}$ in the Einstein frame.

% ------------------------------------------------------------
\section{Gauge Invariance of the Effective Action}
\label{app:gauge}

Under $B_\mu \to B_\mu - \partial_\mu\xi$:
\begin{itemize}
  \item $H_{\mu\nu}$ is manifestly invariant.
  \item The coupling $\Gamma B_\mu J_{\inf}^\mu \to \Gamma B_\mu J_{\inf}^\mu
        - \Gamma (\partial_\mu\xi) J_{\inf}^\mu = \Gamma B_\mu J_{\inf}^\mu
        + \Gamma\xi(\nabla_\mu J_{\inf}^\mu) = \Gamma B_\mu J_{\inf}^\mu$,
        using \eqref{eq:info-conservation}.
\end{itemize}
Hence $\Seff$ is gauge-invariant.

% ------------------------------------------------------------
\section{Computational Pseudocode}
\label{app:code}

The numerical pipeline for field evolution is:
\begin{enumerate}
  \item Initialise $\gm^{(0)}$, $B_\mu^{(0)}$, $\phi^{(0)}$, $J_{\inf}^{(0)}$.
  \item At each time step $n$:
    \begin{enumerate}
      \item Compute $H_{\mu\nu}^{(n)}$ from \eqref{eq:field-strength}.
      \item Solve \eqref{eq:WP} for $B_\mu^{(n+1)}$ (finite-difference
            divergence of field strength on left-hand side).
      \item Solve \eqref{eq:modified-einstein} for $\gm^{(n+1)}$ (ADM
            or BSSN decomposition).
      \item Advance $J_{\inf}^{(n+1)}$ via conservation \eqref{eq:info-conservation}.
      \item Record $\sigmaS^{(n)} = B_\mu^{(n)} J_{\inf}^{\mu,(n)} \ge 0$.
    \end{enumerate}
  \item Output field configurations and entropy production history.
\end{enumerate}

% ============================================================
\section*{Acknowledgements}

The author thanks the open-source scientific community and the developers of
\textsc{SageMath}, \textsc{xAct}, and \textsc{NumPy}/\textsc{SciPy} for tools
used in verification.  This work is dedicated to the Defensive Public Commons.

% ============================================================
\bibliography{references}

\end{document}
